---
title: "Project 1"
author: "Divya Shridar"
date: "10/17/2022"
output: html_document
---

# Develop an Analysis Plan

## Import dataset
```{r Import dataset}
library(NHANES)
head(NHANES)
nrow(NHANES)
```

## Exploratory Data Analysis
```{r}
colnames(NHANES)
```

```{r Examining NAs}
sum(is.na(NHANES$TVHrsDay))
sum(is.na(NHANES$CompHrsDay))
sum(is.na(NHANES$DaysPhysHlthBad))
prop.table(table(NHANES$DaysPhysHlthBad))
2468/10000
```

```{r Race Proportions}
prop.table(table(NHANES$Race1))
```

```{r Sex Orientation}
prop.table(table(NHANES$SexOrientation))
```

```{r Poverty}
prop.table(table(NHANES$Poverty))
```

```{r}
library(ggplot2)
ggplot(NHANES,
       aes(x = Age,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()

# age, gender, outcomes, exposure variables
# population demographics: urban? working? locations?
  # age (continuous): median, IQR
  # percentage for categories
```

```{r}
ggplot(NHANES,
       aes(x = Poverty,
           y = BPSysAve,
           color = Age)) +
  geom_point()
```

```{r}
ggplot(NHANES,
       aes(x = PhysActive,
           y = BPSysAve,
           color = Gender)) +
  geom_point(size = 0.5)
```

```{r}
#table(NHANES$PregnantNow)
NHANESPregnant <- subset(NHANES,
       PregnantNow == 'Yes')
ggplot(NHANESPregnant,
       aes(x = BMI,
           y = BPSysAve)) +
  geom_point()
```

## Inclusion and Exclusion Criteria
```{r Inclusion and Exclusion}
# Adjust for inclusion and exclusion criteria
NHANES_Inclusion <- subset(subset(NHANES,subset = Age >= 18), PregnantNow!='Yes' | is.na(PregnantNow)==TRUE)
NHANES_Inclusion
```
## Aim 1: Two Sample T-test

```{r Poverty using 1}
NHANES_Poor <- subset(NHANES_Inclusion,
                      subset = Poverty < 1)
NHANES_Poor
nrow(NHANES_Poor)

NHANES_NotPoor <- subset(NHANES_Inclusion,
                      subset = Poverty >= 1)
nrow(NHANES_NotPoor)

var(na.omit(NHANES_Poor$BPSysAve)) # 318.0218
var(na.omit(NHANES_NotPoor$BPSysAve)) # 284.7876

t.test(NHANES_Poor$BPSysAve, # var considered not equal
       NHANES_NotPoor$BPSysAve)
```


```{r Socioeconomic Status}
NHANES_Poor <- subset(NHANES_Inclusion,
                      subset = Poverty < 2)
NHANES_Poor
nrow(NHANES_Poor)

NHANES_NotPoor <- subset(NHANES_Inclusion,
                      subset = Poverty >= 2)
nrow(NHANES_NotPoor)

var(na.omit(NHANES_Poor$BPSysAve)) # 318.0218
var(na.omit(NHANES_NotPoor$BPSysAve)) # 284.7876

t.test(NHANES_Poor$BPSysAve, # var considered not equal
       NHANES_NotPoor$BPSysAve)
```
```{r Row Count}
2567+4282
```

## Aim 2: Physical Activity on Blood Pressure
### Apply inclusion and exclusion criteria
```{r}
NHANES_Aim2 <- subset(NHANES_Inclusion, 
                      subset = Age<=75)
nrow(NHANES_Aim2)
```

### Graphics of possible covariates

```{r Age}
ggplot(NHANES_Aim2,
       aes(x = Age,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()
```

```{r Gender}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Gender))
```


```{r BMI}
ggplot(NHANES_Aim2,
       aes(x = BMI,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()
```
```{r Race1}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Race1))
```

```{r}
ggplot(NHANES_Inclusion,
       aes(x = Race1,
           fill = PhysActive)) +
  geom_bar(position = 'dodge')
```


```{r}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Education))
```

### covariates: Age, Gender, Education, Race1, BMI, Depression, TVHrsDay
```{r}
fit2a <- lm(BPSysAve ~ PhysActive*Age, 
          data = NHANES_Aim2)
summary(fit2a)
```

```{r}
fit2 <- lm(BPSysAve ~ PhysActive, 
          data = NHANES_Aim2)
summary(fit2)
```

```{r model fitting} 
fit <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI+Depressed+TVHrsDay, 
          data = NHANES_Aim2)
summary(fit)

# CI for estimates
confint(fit)
```

```{r}
fit3 <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI+Depressed, 
          data = NHANES_Aim2)
summary(fit3)
```

```{r}
fit4 <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit4)
```

```{r}
fit5 <- lm(BPSysAve ~ PhysActive+Age+Gender+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit5)
```


```{r}
fit6 <- lm(BPSysAve ~ PhysActive*Age*Gender*Race1*BMI, 
          data = NHANES_Aim2)
summary(fit6)
```

```{r Assumptions}
plot(fit2, 2)
plot(fit, 2)
plot(fit5, 2)
```


# consider a transformation but not ultimate priority right now

```{r}
fit8 <- lm(log(BPSysAve) ~ PhysActive, 
          data = NHANES_Aim2)
summary(fit8)
```


```{r}
fit7 <- glm(BPSysAve ~ PhysActive*Age*Gender*Race1*BMI, 
          data = NHANES_Aim2)
summary(fit7)

# CI for estimates
confint(fit)
```


