---
title: "Project 1"
author: "Divya Shridar"
date: "10/17/2022"
output: html_document
---

# Develop an Analysis Plan

## Import dataset
```{r Import dataset}
library(NHANES)
head(NHANES)
nrow(NHANES)
```

## Exploratory Data Analysis
```{r}
colnames(NHANES)
```

```{r Examining NAs}
sum(is.na(NHANES$TVHrsDay))
sum(is.na(NHANES$CompHrsDay))
sum(is.na(NHANES$DaysPhysHlthBad))
prop.table(table(NHANES$DaysPhysHlthBad))
2468/10000
```

```{r Race Proportions}
prop.table(table(NHANES$Race1))
```

```{r Sex Orientation}
prop.table(table(NHANES$SexOrientation))
```

```{r Poverty}
prop.table(table(NHANES$Poverty))
```

## Plots for Visualisation
```{r Age, PhysActive}
library(ggplot2)
ggplot(NHANES,
       aes(x = Age,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()
```

```{r Poverty, Age}
ggplot(NHANES,
       aes(x = Poverty,
           y = BPSysAve,
           color = Age)) +
  geom_point()
```

```{r PhysActive, Gender}
ggplot(NHANES,
       aes(x = PhysActive,
           y = BPSysAve,
           color = Gender)) +
  geom_point(size = 0.5)
```

```{r Pregnant only}
#table(NHANES$PregnantNow)
NHANESPregnant <- subset(NHANES,
       PregnantNow == 'Yes')
ggplot(NHANESPregnant,
       aes(x = BMI,
           y = BPSysAve)) +
  geom_point()
```
# START
## Inclusion and Exclusion Criteria
```{r Inclusion and Exclusion}
# Adjust for inclusion and exclusion criteria
NHANES_Inclusion <- subset(subset(NHANES,subset = Age >= 18), PregnantNow!='Yes' | is.na(PregnantNow)==TRUE)
NHANES_Inclusion
```

## Aim 1: Two Sample T-test
```{r Poverty using 1}
NHANES_Poor <- subset(NHANES_Inclusion,
                      subset = Poverty < 1)
NHANES_Poor
nrow(NHANES_Poor)

NHANES_NotPoor <- subset(NHANES_Inclusion,
                      subset = Poverty >= 1)
nrow(NHANES_NotPoor)

var(na.omit(NHANES_Poor$BPSysAve)) # 318.0218
var(na.omit(NHANES_NotPoor$BPSysAve)) # 284.7876

t.test(NHANES_Poor$BPSysAve, # var considered not equal
       NHANES_NotPoor$BPSysAve)

# included plot
ggplot(data = NHANES_Inclusion,
       mapping = aes(x = Poverty,
                     y = BPSysAve,
                     color = Poverty<1)) +
  geom_point()
```


```{r Socioeconomic Status}
NHANES_Poor <- subset(NHANES_Inclusion,
                      subset = Poverty < 2)
NHANES_Poor
nrow(NHANES_Poor)

NHANES_NotPoor <- subset(NHANES_Inclusion,
                      subset = Poverty >= 2)
nrow(NHANES_NotPoor)

var(na.omit(NHANES_Poor$BPSysAve)) # 318.0218
var(na.omit(NHANES_NotPoor$BPSysAve)) # 284.7876

t.test(NHANES_Poor$BPSysAve, # var considered not equal
       NHANES_NotPoor$BPSysAve)

# included plot
ggplot(data = NHANES_Inclusion,
       mapping = aes(x = Poverty,
                     y = BPSysAve,
                     color = Poverty<2)) +
  geom_point()
```
```{r Row Count}
2567+4282
```

## Aim 2: Physical Activity on Blood Pressure
```{r Inclusion and Exclusion}
NHANES_Aim2 <- subset(NHANES_Inclusion, 
                      subset = Age<=75)
nrow(NHANES_Aim2)
```

### Graphics of possible covariates
```{r Age}
ggplot(NHANES_Aim2,
       aes(x = Age,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()
```

```{r Gender}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Gender))
```


```{r BMI}
ggplot(NHANES_Aim2,
       aes(x = BMI,
           y = BPSysAve,
           color = PhysActive)) +
  geom_point()
```

```{r Race1}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Race1))
```

```{r Race1 bar plot}
ggplot(NHANES_Inclusion,
       aes(x = Race1,
           fill = PhysActive)) +
  geom_bar(position = 'dodge')
```


```{r Education Facet}
ggplot(NHANES_Aim2,
       aes(x = PhysActive,
           y = BPSysAve)) +
  geom_point() +
  geom_boxplot() +
  facet_grid(cols = vars(Education))
```
Chosen covariates: Age, Gender, Education, Race1, BMI, Depression, TVHrsDay

### Model Fitting
```{r fit without covariates}
fit2 <- lm(BPSysAve ~ PhysActive, 
          data = NHANES_Aim2)
summary(fit2)
```

```{r model fitting with covariates w/o interaction} 
fit <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI+Depressed+TVHrsDay, 
          data = NHANES_Aim2)
summary(fit)

# CI for estimates
confint(fit)
```

```{r remove tvhrsperday}
fit3 <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI+Depressed, 
          data = NHANES_Aim2)
summary(fit3)
```

```{r remove depressed}
fit4 <- lm(BPSysAve ~ PhysActive+Age+Gender+Education+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit4)
```

```{r remove education}
fit5 <- lm(BPSysAve ~ PhysActive+Age+Gender+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit5)
```

```{r remove physactive}
fit5a <- lm(BPSysAve ~ Age+Gender+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit5a) # performs exactly the same as with physactive
```

```{r with interaction}
fit6 <- lm(BPSysAve ~ PhysActive*Age*Gender*Race1*BMI, 
          data = NHANES_Aim2)
summary(fit6)
```

```{r interaction without physactive}
fit6a <- lm(BPSysAve ~ Age*Gender*Race1*BMI, 
          data = NHANES_Aim2)
summary(fit6a) # performs worse than with physactive when looking at interaction
```


```{r second best: log transform}
fit8 <- lm(log(BPSysAve) ~ PhysActive+Age+Gender+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit8) # 0.1894

# improvement from 0.1848
confint(fit8)
```

```{r log transform without phys}
fit8 <- lm(log(BPSysAve) ~ Age+Gender+Race1+BMI, 
          data = NHANES_Aim2)
summary(fit8) # same result as with phys
```

```{r best: log and interaction}
fit7 <- lm(log(BPSysAve) ~ PhysActive*Age*Gender*Race1*BMI, 
          data = NHANES_Aim2)
summary(fit7) # 0.2299, 0.0099 better than without Physactive

confint(fit7)
```


```{r just physactive and age}
fit2a <- lm(BPSysAve ~ PhysActive*Age, 
          data = NHANES_Aim2)
summary(fit2a)
```

```{r Check Assumptions}
# 1. check for linear relationship between x and y
  # smoothes trend is very stable
plot(fit2, 1) # just x and y
plot(fit, 1) # x and y with all covariates
plot(fit7, 1) # best perfroming model
plot(fit8, 1) # second best perfroming model

# 2. independent observations due to the way the data was collected

# 3. residuals are normally distributed
  # looks reasonably good
plot(fit2, 2) # just x and y
plot(fit, 2) # x and y with all covariates
plot(fit7, 2) # best perfroming model
plot(fit8, 2) # second best perfroming model

# 4. constant variance of the results
  # no evidence of increasing or decreawsing spread of range of fitted vals
plot(fit2, 1) # just x and y
plot(fit, 1) # x and y with all covariates
plot(fit7, 1) # best perfroming model
plot(fit8, 1) # second best perfroming model
```

## Deliverables
* Include table of distribution of: age, gender, outcomes, exposure variables
  * population demographics: urban? working? locations?
  * age (continuous): median, IQR
  * percentage for categorical variables 
* Assumptions before testing
* Confint of estimates

